---
title: "Vaccination Rates LA County"
author: "Oscar Monroy"
date: "10/30/2021"
output: pdf_document
---

# Step 1: Importing and Fixing the Data

### Variable name on R vs. Variable name on Excel:

### X12..with.1..Dose = 12+ with 1+ Dose
### Population..12.. = Population (12+)
### X12..Pop..Vaccinated.... = 12+ Pop. Vaccinated (%)

```{r}
la_c <- read.csv("LA-County-COVID-19-Vaccine-Dashboard-Community-2021-10-24.csv")
head(la_c) # Looks messy, for R standards
dim(la_c) # Dimeensions of our data frame
la_c$City.Community <- as.character(la_c$City.Community)
# For the next 3 lines, we'll change the class of 
# the last 3 variables to numeric to make things easier.
la_c$X12..with.1..Dose <- as.numeric(gsub("[\\,]", "", la_c$X12..with.1..Dose))
la_c$Population..12.. <- as.numeric(gsub("[\\,]", "", la_c$Population..12..))
la_c$X12..Pop..Vaccinated.... <- as.numeric(gsub("[\\%,]", "", la_c$X12..Pop..Vaccinated....))
head(la_c) # Looks much better now
```

# Step 2: Cleaning the Data

```{r}
summary(la_c)
# Looks about right considering this is LA County,
# however, those NA's need to be taken care of.
nr <- which(is.na(la_c$X12..with.1..Dose) == TRUE) # Checks index of NA's
nr
which(is.na(la_c$Population..12..) == TRUE)
which(is.na(la_c$X12..Pop..Vaccinated....) == TRUE)
# Appears that all our NA's happen on the same observations...
la_c2 <- la_c[-nr, ]
summary(la_c2) # No NA's; success
# Now we check for duplicate observations
which(duplicated(la_c2) == TRUE) # No duplicates
# We are done with cleaning the data
```

# Step 3: Visualizations and Analysis

```{r}
la_c3 <- la_c2
# Let's compare the population to the percent of that 
# population that vaccinated.
plot(la_c3$X12..Pop..Vaccinated.... ~ la_c3$Population..12.., 
     xlab = "Population (12+)", 
     ylab = "Percent Vaccinated (12+)",
     main = "Total Population vs Percent Vaccinated (Age 12+)") # Doesn't show a pattern...

# Let's see if the more populated areas of 
# LA County have a higher vaccination rate.
# We'll use a population of 15k as our arbitrary cutoff point
# between sparsely populated and highly populated.
tp <- which(la_c3$Population..12.. >= 15000)
over10k <- rep("N", 324)
over10k[tp] <- "Y"
la_c3 <- cbind(la_c3, over10k)
table(la_c3$over10k) # Amount of areas over/under pop. of 15k
plot(la_c3$over10k, la_c3$X12..Pop..Vaccinated...., 
     xlab = "Pop. of 12+ Over 15k?", 
     ylab = "Percent of Pop. Vaccinated (12+)",
     main = "Higher Population vs Lower Population Vax Rates")

par(mfrow = c(1, 2))
plot(la_c3$X12..Pop..Vaccinated....[tp] ~ la_c3$Population..12..[tp],
     ylim = c(0, 100), xlab = "Popuation <= 15k (12+)",
     ylab = "Percent Vaccinated (12+)", col = "green", 
     main = "Pop. <= 15k vs % Vaccinated")
abline(h = mean(la_c3$X12..Pop..Vaccinated....[tp]), col = "purple")
plot(la_c3$X12..Pop..Vaccinated....[-tp] ~ la_c3$Population..12..[-tp], 
     ylim = c(0, 100), xlab = "Popuation > 15k (12+)",
     ylab = "Percent Vaccinated (12+)", col = "red", 
     main = "Pop. > 15k vs % Vaccinated")
abline(h = mean(la_c3$X12..Pop..Vaccinated....[-tp]), col = "purple")
legend("bottomright", legend = c("mean %"), col = c("purple"), lty = 1)
```

### We do see that the more populated areas have a higher vaccination rate
### on average as we can see from the scatterplot where the higher pop. areas
### have a high concentration between 60% and 90%, whereas the lower pop. 
### areas seems to show a more varied spread of the percentage vaccinated
### and a slightly lower average in general. 

```{r}
# Now let's a try a linear regression model
la_m <- lm(la_c3$X12..Pop..Vaccinated.... ~ la_c3$Population..12..) 
summary(la_m)
plot(la_c3$X12..Pop..Vaccinated.... ~ la_c3$Population..12.., 
     xlab = "Population (12+)", 
     ylab = "Percent Vaccinated (12+)",
     main = "Total Population vs Percent Vaccinated (Age 12+)")
abline(la_m, col = "red")
```

### While the summary of the model seems to be promising with how it
### shows a positive trend in vax rate as pop. increases, adding a 
### regression line to the first scatterplot doesn't really work,
### although it could probably be useful to predict the vaccination rates
### of other districts not listed here. Unfortunately, this data set 
### is fair;y barebones, so we can't really do much else.
### I'd love to try and add political alignments for each districts, but 
### unfortunately, there is no current data set with such info
### and if it does, it groups all unincorporated regions into one. 
### For now, we do see that higher population areas do make a difference in 
### vaccination rates, at least in the LA County. 
